---
title: Interprovincial Survey of DST
author: 'Ting Hsu, Samantha Hassal'
date: '2020-10-06'
slug: ps2
categories: []
tags: []
output:
  blogdown::html_page:
    toc: true
    toc-depth: 1
---


```{r setup, include = FALSE}
#title􏰃 date􏰃 authorship􏰃 non􏰁technical executive summary􏰃 introduction􏰃 survey methodology􏰃 results􏰃 discussion􏰃 appendices that detail the survey􏰃 and references

library(readr)
library(dplyr)
library(ggplot2)
library(tidyverse)
```
*Authors: S. Hassal, T. Hsu*

# Non-technical Executive Summary

In Ontario and Saskatchewan, people more often than not find themselves tired and unproductive under DST. The amount of hours that people want to sleep is not affected by the prescence of the DST system.


# Introduction

This survey seeks to determine the effects of Daylight Saving Time policy on the 
population and their respective opinions on the subject. Which will aid the Conservative Party's decision-making on whether or not DST should be kept. To clarify, Day Light Saving time is where on two days of the year we shift the clock forward one hour, and later backwards one hour. The original purpose of this is to make more use out of daytime during the summer months, as it moves one hour of daylight in the morning to the evening.

In this report, the data is gathered with an online form targeting the populations of Saskatchewan and Ontario. These two provinces are chosen because one uses DST (Ontario), and the other doesn't (Saskatchewan). However, Saskatchewan is _always_ on DST. Meaning that the time zone for Saskatchewan should be Mountain Time Zone (GST -7:00), but it is using Central Standard Time (GST -6:00).

# Survey Methodology
```{r, include = FALSE}
# Questions REFERENCE
# Q1 Do you have a sleep disorder?
# Q2 Closest Major City
# Q3 During Daylight Savings, everyday inconveniences make me upset
# Q4 During Daylight Savings, I struggle to put my thoughts in words.
# Q5 When I don't get enough sleep, I feel my performance at work or school suffer.
# Q6 I find it harder to get to bed on time during daylight saving
# Q7 During daylight savings, I catch myself waking up in the middle of the night because I am worried
# Q8 How much sleep is needed to function (Number in Hours)
```


The data of this report is collected online through Google Forms, it seeks to compare two distinctly different provinces in terms of DST: Saskatchewan and Ontario, and its resident's opinion on the matter. The reason for this is that we want to have two independent groups to compare.

We chose (relatively) major cities that are in the southern parts of their respective provinces, as shown below. These cities share similar latitudes and therefore should have similar hours of daylight (Not factoring weather). 

Latitude of Cities (In Degrees North):

Ontario Cities:    

  - Windsor: 42
  - Toronto: 43 
  - Sarnia: 42
  - Ottawa: 45
  - Kingston: 44
  - Thunder Bay: 48
  
Saskatchewan Cities:

  - Saskatoon: 52
  - Regina: 50 
  - Timmins: 48
  - Uranium City: 59
  

We are interested in the following information:

 - Functionality of life
 - Opinion on DST (Positive of Negative)
 
Where the functionality is determined by asking a series of questions on their sleeping quality, which is what we assumed to be a big factor. On the other hand, opinions about the DST is presented as whether or not people think that they are more irritable or less able to think while DST system is in place. There are multiple survey questions regarding these so that we can get a lower error. 

Since we are basing most of the data on sleeping habits, we will not consider anyone with sleeping conditions. The first question of the survey is designed so that this issue is addressed.


# Data

The data is randomly generated and sampled with the following assumptions:

  - Only 0.2 % of the target population responded.
  Average duration of sleep 9.5 hours and standard deviation 1.7 hours
  - Small percentage of the target audience completed the survey
  - Equal likelihood that people within a specific region say they live nearest to a        given city (eg, people who live in Eastern Ontario are just as likely to say they       live near Kingston as they are to live near Ottawa)
  - Vast majority (~85%) of people agreed they can’t get to sleep during daylight
    savings time
  - People in Saskatchewan still knows what DST feels like, even though they do not         currently use DST.

```{r graphs, echo = FALSE}
#constants - city/town population (Sask)
CITIES <- 655313
DIV_18 <- 37064

#constants - city/town population (Ont) 
#the figures for ontario came from a few quiick google searches based on region, the figures for sask came from the population report on the census
GTA <- 5.928*10^6
SW <- 2.584*10^6
EAST <- 1.763*10^6
NW <- 231691

#Total population of areas of interest
TOT <- GTA+SW+EAST+NW+DIV_18+CITIES

#probability vector population
pop_vec <- 1/TOT * c(GTA,SW, SW,EAST, NW, CITIES, CITIES, EAST,NW,DIV_18)

#vector of cities
cities <- c('Toronto', 'Windsor', 
            'Sarnia', 'Ottawa', 'Thunder Bay', 
            'Saskatoon', 'Regina', 'Kingston', 
            'Timmins', 'Uranium City')
cities_ON <- c('Toronto', 'Windsor', 
            'Sarnia', 'Ottawa', 'Thunder Bay', 'Kingston')
cities_SA <- c('Saskatoon', 'Regina', 'Timmins', 'Uranium City')
#agree-disagree vector
agree_dis <- c('Completely disagree','Disagree','Neutral','Agree','Completely agree')

#number of hours of sleep
sleep <- seq(6, 13, by = 0.1)
sleep_prob <- dnorm(sleep, mean = 9.5, sd = 1.7)

#seeding for random variables (assume 0.2% of the total population take survey)
n = 0.002*TOT
set.seed(n)
#samples for survey questions
Q1 = sample(c('yes','no'), n, replace = TRUE, prob = c(0.05, 0.95))
Q2 = sample(cities, n, replace = TRUE, prob=pop_vec)
Q3 = sample(agree_dis, n, replace = TRUE, prob=c(.2,.2,.2,.2,.2))
Q4 = sample(agree_dis, n, replace = TRUE, prob=c(.15,.2,.15,.2,.3))
Q5 = sample(agree_dis, n, replace = TRUE, prob=c(.05,.2,.05,.4,.2))
Q6 = sample(agree_dis, n, replace = TRUE, prob=c(.05,.05,.05,.85,0))
Q7 = sample(agree_dis, n, replace = TRUE, prob=c(0,.15,.7,.15,0))
Q8 = sample(sleep, n, replace = TRUE, prob=sleep_prob)

#collect in a table
results <- tibble(Q1,Q2,Q3,Q4,Q5,Q6,Q7,Q8)

#filter out people with sleep disorders
working <- results %>%
filter(Q1=="no")


#look at who agreed with Question 6
q6_agree <- working %>%
filter(Q6=="Agree")

#make a bar graph of where people lived
bar_tor <- ggplot(q6_agree, aes(x=Q2)) + geom_bar() + labs(x="City", y='number of responses')
bar_tor + coord_flip()

#retrieve hours of sleep for each city
tor <- working %>%
filter(Q2=='Toronto')%>%
select(Q8)

ott <- working %>%
filter(Q2=='Ottawa')%>%
select(Q8)

sask <- working %>%
filter(Q2=='Saskatoon')%>%
select(Q8)

#histogram of hours of sleep
hist(deframe(tor), main="Hours of Sleep per Night (Toronto)",
xlab="Hours",
col="darkmagenta")

hist(deframe(ott), main="Hours of Sleep per Night (Ottawa)",
xlab="Hours",
col="goldenrod")

hist(deframe(sask), main="Hours of Sleep per Night (Saskatoon)",
xlab="Hours",
col="chocolate")


#Bargraph of Q3-Q7 Responses Combined:
# combined<- unite(working,"Response", Q3:Q7, na.rm = TRUE, remove = FALSE)
q3q7Dat <- select(working, Q2:Q7)
q3q7Comb <- pivot_longer(q3q7Dat, Q3:Q7)
q3q7Comb %>% 
count(value) %>% 
mutate(prop = n/sum(n)) %>% 
ggplot(aes(x = value, y = prop, group = value)) +
geom_col(aes(fill = value), position = "dodge") +
geom_text(aes(label = scales::percent(prop), y = prop, group = value)) +
theme(axis.text.x=element_blank(), axis.title.x = element_blank()) +
labs(title = "Overall Proportion of Response Towards DST Being Bad")

q3q7Comb %>% 
filter(Q2 %in% cities_ON) %>%  
count(value) %>% 
mutate(prop = n/sum(n)) %>% 
ggplot(aes(x = value, y = prop, group = value)) +
geom_col(aes(fill = value), position = "dodge") +
geom_text(aes(label = scales::percent(prop), y = prop, group = value)) +
theme(axis.text.x=element_blank(), axis.title.x = element_blank()) +
labs(title = "Proportion of Response Towards DST Being Bad ON")

q3q7Comb %>% 
filter(Q2 %in% cities_SA) %>%  
count(value) %>% 
mutate(prop = n/sum(n)) %>% 
ggplot(aes(x = value, y = prop, group = value)) +
geom_col(aes(fill = value), position = "dodge") +
geom_text(aes(label = scales::percent(prop), y = prop, group = value)) +
theme(axis.text.x=element_blank(), axis.title.x = element_blank()) +
labs(title = "Proportion of Response Towards DST Being Bad SA")
  

#TODO Generate the rest of the responses.


#0.2% of the target audience responded to the survey


```

ON = Ontario, SA = Saskatchewan


*The RMarkdown file for this report can be found at: https://github.com/T-HSU-12345/sta304/tree/master/content/post/2020-10-06-ps2.Rmd*

# Discussion

Toronto:  
mean: `r mean(tor$Q8)` hours, standard deviation: `r sd(tor$Q8)` hours  
Ottawa:  
mean: `r mean(ott$Q8)` hours, standard deviation: `r sd(ott$Q8)` hours  
Saskatchewan:  
mean: `r mean(sask$Q8)` hours, standard deviation: `r sd(sask$Q8)` hours  

The mean is calculate using the following function: 

$\mu = \frac{\sum h}{n}$

Where the nominator $\sum h$ represents the sum of all the hours combined, and the denominator $n$ represents the total number of answers. $\mu$ represents the mean.

The standard deviation is calculated using the following function:

$\mu = \sqrt{\frac{\sum(x_i - \mu)^2}{N}}$

Where the nominator $\sum(x_i - \mu)^2$ is the sum of each input of hours minus the mean all squared, and the denominator $N$ is the total number of participants. This calculates what proportion of people have the similar sleeping schedules. A standard deviation of `r sd(tor$Q8)` hours for Toronto, for example, meant that 68.2 % of people have preferred sleeping time of `r mean(tor$Q8) - sd(tor$Q8)` - `r mean(tor$Q8) + sd(tor$Q8)` hours. Which is the mean plus or minus one standard deviation.

As shown above, the three major cities across both provinces have people following a similar normal distribution curve. With just over half of the people for both provinces have opinions against DST, of which about a quarter are neutral. 

This survey is purposefully kept short, with only 7 simple questions of multiple choice, and one simple numerical question (how many hours of sleep do you need?), this aims to reduce the chance of people randomly selecting the answers due to boredom/lack of knowledge.
There are several factors that we can not account for with this survey. We can only determine who have poor opinion of DST, but not who likes DST, due to the nature of the survey questions. Since the people who answered "Neutral" can swing either way. There are certain inherent inaccuracies, since there might be people who lived in the northern part of the province and are more affected by daylight shift, but are not accounted for withe the choice of nearest city mostly being in the southern part of either provinces.

# Results






### References

- Wickham et al., (2019). Welcome to the tidyverse. Journal of Open Source
  Software, 4(43), 1686, https://doi.org/10.21105/joss.01686

- Hadley Wickham, Romain François, Lionel Henry and Kirill Müller (2020). dplyr:
  A Grammar of Data Manipulation. R package version 1.0.2.
  https://CRAN.R-project.org/package=dplyr

- H. Wickham. ggplot2: Elegant Graphics for Data Analysis. Springer-Verlag New
  York, 2016.

- Hadley Wickham, Jim Hester and Romain Francois (2018). readr: Read Rectangular
  Text Data. R package version 1.3.1. https://CRAN.R-project.org/package=readr

- R Core Team (2020). R: A language and environment for statistical computing. R
  Foundation for Statistical Computing, Vienna, Austria. URL
  https://www.R-project.org/.
  
- JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi and Kevin
  Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and Winston Chang and
  Richard Iannone (2020). rmarkdown: Dynamic Documents for R. R package version
  2.3. URL https://rmarkdown.rstudio.com.

- Yihui Xie (2020). blogdown: Create Blogs and Websites with R Markdown. R
  package version 0.20.

- Yihui Xie, Alison Presmanes Hill, and Amber Thomas (2017). blogdown: Creating
  Websites with R Markdown. Chapman and Hall/CRC. ISBN 978-0815363729






