---
title: "CES: Comparing Opinions of the Economy between the Differently Employed"
author: Ting Hsu
date: '2020-09-22'
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
slug: ps1
---

```{r setup, include=FALSE}
library(devtools)
library(dplyr)
library(tidyverse)
library(cesR)
#install.packages("gridExtra")
library(gridExtra)
library(visdat)
```

# Abstract

Using data from the 2019 Canadian Elections Study, three different questions about the economy and given to people of different employment statuses (Students, Retired, Part Time, Full Time, Unemployed, and Disabled), and their opinions are compared. The data is then organized with the goal to see if there will be major differences in opinion between people of different employment status. It seems that people tend to not think the economy is better, and think it is mostly worse off.

***

# Introduction

Is the Canadian economy in recession? I am not the expert on this subject. However, the Canadian Election Survey have collected information from 37822 people about their opinion, and I can organize their opinions to get a general idea as to what people THINK about the economy. The survey contains a myriad of questions such as employment status, religion, and income. This report looks at how might employment status affect the people's opinion on how well the economy is doing.

There are three sections of the survey that will be compared, and they are: 

* econ_retro: How people felt the economy have been in the past year.
* econ_fed: If people think that the government have made an impact on the economy - whether it is positive or negative.
* econ_self: How did people feel the effects of the government policies economically.  

Through the exploration of data, I found that unemployed and disabled people are more likely, according to this survey, to view the economy to be worse. Though a significant portion of them still think that there are no significant differences. Most importantly, at any given time only a minority in any group think that the economy is getting better.

This is only a preliminary analysis of the data, and no statistical model have been fitted to match the data. Mainly because the study is done on categorical data with only four different choices. Which means that fitting a statistical model is an improvement that can be made, and of course, a weakness in itself also. Regardless, another possible area that can be explored is other factors that may affect people's opinion on the economy.

```{r download, warning=FALSE, message=FALSE, echo = FALSE, include = FALSE}
# install.packages("devtools")
# devtools::install_github("hodgettsp/cesR")
# Get the CES data
cesR::get_decon()
```

# Data

```{r functions, echo = FALSE}

#This is a totaling function, cat is the column in decon, target is the string to sort the input wanted from cat, and type is the sub group in the cat. Returns the total rows of target population is type is NULL. Otherwise, returns the amount of people that did not give no opinion.
#e.g: If you want to find the amount of Students that do not have an opinion on econ_retro: total(decon$employment, c("Student", "Student and working for Pay"), decon$econ_retro)

total <- function(cat, target, type){
  if (is.null(type)){
    return(nrow(filter(decon, cat %in% target)))
  } else {
    return(nrow(filter(decon, cat %in% target)) - nrow(filter(decon, cat %in% target & type %in% c("Don't know/ Prefer not to answer"))))
  }
}
```
Of `r nrow(decon)` people, 2456 of them did not fill in the data and therefore are not considered as part of the following graphs. In the survey, the students are split into Students and Students with Pay, but they are combined in this report as simply "Student".

Here are the general numbers:

- Total amount of Students: `r total(decon$employment, c("Student", "Student and working for Pay"), NULL)`
- Total amount of Part Timers: `r total(decon$employment, c("Working for pay part-time"), NULL)`
- Total amount of Full Timers: : `r total(decon$employment, c("Working for pay full-time"), NULL)`
- Total amount of Retired: `r total(decon$employment, c("Retired"), NULL)`
- Total amount of Unemployed: `r total(decon$employment, c("Unemployed/ looking for work"), NULL)`
- Total amount of Disabled: `r total(decon$employment, c("Disabled"), NULL)`
- Total amount of Self Employed: `r total(decon$employment, c("Self employed (with or without employees)"), NULL)`

### People's View of the Economy over the Past Year
_Figures are numbered from top to bottom, and left to right. As in, Student's Opinion -> Part Time -> Full Time -> Retired -> Unemployed -> Disabled -> Self Employed_  

_Fig. 1 ~ 7_

```{r graph-econ_ER, echo = FALSE}
#Displaying Bar Graph of various employment groups of their opinions on Canada's economy within the past year

#Takes optF as the vector that act as the filter input, topic is the title,
#graphs the percentage graph of econ_retro

graphER<- function( optF, topic){
  decon %>%
  filter(decon$employment %in% optF) %>% 
  count(econ_retro) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_retro, y = prop, group = econ_retro)) +
  geom_col(aes(fill = econ_retro), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_retro)) +
  theme(legend.position = "none", axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = topic) -> erGraph
  return(erGraph)
}

#Display first graph with Legend (All the graphs share the same Legend)
decon %>%
  filter(decon$employment %in% c("Student", "Student and working for Pay")) %>% 
  count(econ_retro) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_retro, y = prop, group = econ_retro)) +
  geom_col(aes(fill = econ_retro), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_retro)) +
  theme(axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = "Student's Opionion")

graphER( c("Working for pay part-time"), 
         "Part Time") -> pt

graphER(c("Working for pay full-time"), "Full Time") -> ft

graphER(c("Retired"), "Retired") -> re

graphER(c("Unemployed/ looking for work"), "Unemployed") -> um

graphER(c("Disabled"), "Disabled")-> ds

graphER(c("Self employed (with or without employees)"), "Self Employed") -> se

grid.arrange(pt, ft, re, um, ds, se, ncol=2, nrow =3)
```

- The average difference between "Got Worse" and "Got Better": `r ((32.4 - 11.6) + (37.1 - 12) + (37 - 13.2) + (34 - 17) + (46.5 - 9.2) + (46.0 - 9.4) + (38.9 - 12.3)) / 7`%
 

### People's View of the Government's Work on the Economy
_Fig. 8 ~ 13_

```{r graph-econ_EF, echo = FALSE}
#Displaying graph of how the various employment groups view the government

#Takes optF as the vector that act as the filter input, topic is the title,
#graphs the percentage graph of econ_fed

graphEF<- function( optF, topic){
  decon %>%
  filter(decon$employment %in% optF) %>% 
  count(econ_fed) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_fed, y = prop, group = econ_fed)) +
  geom_col(aes(fill = econ_fed), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_fed)) +
  theme(legend.position = "none", axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = topic) -> erGraph
  return(erGraph)
}

#Display first graph with Legend (All the graphs share the same Legend)
decon %>%
  filter(decon$employment %in% c("Student", "Student and working for Pay")) %>% 
  count(econ_fed) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_fed, y = prop, group = econ_fed)) +
  geom_col(aes(fill = econ_fed), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_fed)) +
  theme(axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = "Student's Opionion")

#Print the rest of the graphs
graphEF( c("Working for pay part-time"), "Part Time") -> ptf

graphEF(c("Working for pay full-time"), "Full Time") -> ftf

graphEF(c("Retired"), "Retired") -> rtf

graphEF(c("Unemployed/ looking for work"), "Unemployed") -> umf

graphEF(c("Disabled"), "Disabled")-> dsf

graphEF(c("Self employed (with or without employees)"), "Self Employed") -> sef

#Arrange the rest of the graphs
grid.arrange(ptf, ftf, rtf, umf, dsf, sef, ncol=2, nrow =3)
```

- The average difference between Worse and Better: `r ((28.6 - 16.5) + (36 - 15.1) + (37.7- 17.61) + (33.9 - 19.7) + (40.2 - 13.7) + (41.8 - 13.6) + (39.49 - 14.02)) / 7`%

### Government's Policies' Effect on People
_Fig. 14 ~ 19_

```{r graph-econ_ES, echo = FALSE}
#Displaying graph of how the various employment groups view their situation as compared to before

#Takes optF as the vector that act as the filter input, topic is the title,
#graphs the percentage graph of econ_self

graphES<- function( optF, topic){
  decon %>%
  filter(decon$employment %in% optF) %>% 
  count(econ_self) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_self, y = prop, group = econ_self)) +
  geom_col(aes(fill = econ_self), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_self)) +
  theme(legend.position = "none", axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = topic) -> erGraph
  return(erGraph)
}

#Display first graph with Legend (All the graphs share the same Legend)
decon %>%
  filter(decon$employment %in% c("Student", "Student and working for Pay")) %>% 
  count(econ_self) %>% 
  mutate(prop = n/sum(n)) %>% 
  ggplot(aes(x = econ_self, y = prop, group = econ_self)) +
  geom_col(aes(fill = econ_self), position = "dodge") +
  geom_text(aes(label = scales::percent(prop), y = prop, group = econ_self)) +
  theme(axis.text.x=element_blank(),
  axis.ticks.x=element_blank(), axis.title.x = element_blank()) +
  labs(title = "Student's Opionion")

graphES( c("Working for pay part-time"), "Part Time") -> pts

graphES(c("Working for pay full-time"), "Full Time") -> fts

graphES(c("Retired"), "Retired") -> rts

graphES(c("Unemployed/ looking for work"), "Unemployed") -> ums

graphES(c("Disabled"), "Disabled")-> dss

graphES(c("Self employed (with or without employees)"), "Self Employed") -> ses

grid.arrange(pts, fts, rts, ums, dss, ses, ncol=2, nrow =3)
```

- The average difference between "Better" and "Worse": `r ((31.5 - 12.6 ) + (35.9 - 10.5) + (37.3 - 11.9) + (30.7 - 7.2)  + (46.8 - 7.9) + (44.6 - 7.9) + (37.8 - 9.3)) / 7`%

# Discussion
The main theme across Fig. 1-7 are that everyone except for the unemployed and the disabled people share similar opinions of economy, with roughly 10% of Disabled and Unemployed people are shifted towards "Got Worse" as compared to "Stayed about the same.  

Across Fig. 8-13 the average difference between "Stayed about the same" and "Worse" is about the same, but still there are much more people with a negative outlook on the economical policies than positive outlook. 

Across Fig. 14 ~ 19, it has the biggest average difference between Worse and Better, and it has approximately the same results as Fig. 1 ~ 7, as government's policy and economy seem to go hand in hand.

The surprising thing is that across all three economical opinions, there is always more than a third of any group that think the economy has not shifted at all, and it is almost unanimous in the fact that a big proportion of the economy is getting worse. Though the students have the most variability in terms of % of people thinking that the economy is better.

This survey is conducted online, and it has many columns containing non-responses. None responses are shown as a gray column in the figure below. However, the columns used are the ones that are mostly in green, meaning that there's a response.

```{r see_none_response, echo = FALSE}
visdat::vis_dat(decon)
```



## Weaknesses and next steps

The statistical means calculated in the data sections are not weighted, which meant that the smaller populations like unemployed have the same weight in calculating the average as the full time workers, which make up almost half of all the data point. Also, as mentioned before there are no statistical models fitted to this report, and the analysis done is mostly rudimentary.

Possible next steps can be to dig further into the data and:

- Put a weight on each employment area based on amount of people
- Find another variable that may tie employment type to their opinion
- Fit a model where possible



# References

- Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Online Survey', https://doi.org/10.7910/DVN/DUS88V, Harvard Dataverse, V1
- Paul A. Hodgetts and Rohan Alexander (2020). cesR: Access the CES Datasets a Little Easier.. R package version 0.1.0.
-  Hadley Wickham, Jim Hester and Winston Chang (2020). devtools: Tools to Make Developing R Packages Easier. https://devtools.r-lib.org/,https://github.com/r-lib/devtools.
- R Core Team (2020). R: A language and environment for statistical computing. R, Foundation for Statistical Computing, Vienna, Austria. URL, https://www.R-project.org/.
- JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi and Kevin Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and Winston Chang and Richard Iannone (2020). rmarkdown: Dynamic Documents for R. R package version 2.3. URL https://rmarkdown.rstudio.com.
- Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Online Survey',
           https://doi.org/10.7910/DVN/DUS88V, Harvard Dataverse, V1
- Baptiste Auguie (2017). gridExtra: Miscellaneous Functions for "Grid" Graphics. R package version 2.3.
  https://CRAN.R-project.org/package=gridExtra
  




