---
title: PS1
author: Ting Hsu
date: '2020-09-22'
categories:
  - R
tags:
  - plot
  - R Markdown
  - regression
slug: ps1
---

```{r setup, include=FALSE}
library(devtools)
library(dplyr)
library(tidyverse)
library(cesR)
library(visdat)
library(cesR)
library(skimr)
```

# Introduction

Three paragraph statement.

First paragraph: General. Motivational. Majors specific finding. 

Second paragraph: More about what you did and found.

Third paragraph: Outline of the paper. Future work. Weaknesses.


```{r download, warning=FALSE, message=FALSE, echo = FALSE, include = FALSE}
# install.packages("devtools")
# devtools::install_github("hodgettsp/cesR")

cesR::get_decon()
```

# Data

The CES dataset is....

```{r}
head(decon)
```
Some initial explorations that would be used to guide this section.

```{r}
#Individual Territories
deconNW <- filter(decon, decon$province_territory == "Northwest Territories")
deconYU <- filter(decon, decon$province_territory == "Yukon")
deconNU <- filter(decon, decon$province_territory == "Nunavut")
#All Provincial Data
`%!in%` = Negate(`%in%`)
deconPRO <-filter(decon, decon$province_territory %!in% c("Northwest Territories",
                                                          "Yukon", "Nunavut") )
```


The variables we focus on are....
Education

Summary statistics for these variables...

You must include graphs.


```{r}
#Showing Education Levels between provinces
# decon %>% 
#   ggplot(aes(x = education, fill = education)) +
#   geom_bar() + theme(legend.position = "top", axis.text.y=element_blank()) + guides(fill = "legend") + scale_fill_manual(values = c('#e6194b', '#3cb44b', '#ffe119', '#4363d8', '#f58231', '#911eb4', '#46f0f0', '#f032e6', '#bcf60c', '#fabebe', '#008080', '#e6beff', '#9a6324', '#fffac8')) + facet_wrap(~ decon$province_territory) + coord_flip()
decon %>% 
  
                                                                                                                          
```


```{r}
decon %>% 
  ggplot(aes(x = income)) +
  geom_histogram()
```


In this graph we find that the distribution of income is heavily skewed....




# Discussion

My discussion... main findings of your graph.


Features of your survey.... Non-response... etc. And how these could affect the results....




## Weaknesses and next steps

My weaknesses...






# References

- Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Online Survey', https://doi.org/10.7910/DVN/DUS88V, Harvard Dataverse, V1
- Paul A. Hodgetts and Rohan Alexander (2020). cesR: Access the CES Datasets a Little Easier.. R package version 0.1.0.
-  Hadley Wickham, Jim Hester and Winston Chang (2020). devtools: Tools to Make Developing R Packages Easier. https://devtools.r-lib.org/,https://github.com/r-lib/devtools.
- R Core Team (2020). R: A language and environment for statistical computing. R, Foundation for Statistical Computing, Vienna, Austria. URL, https://www.R-project.org/.
- JJ Allaire and Yihui Xie and Jonathan McPherson and Javier Luraschi and Kevin Ushey and Aron Atkins and Hadley Wickham and Joe Cheng and Winston Chang and Richard Iannone (2020). rmarkdown: Dynamic Documents for R. R package version 2.3. URL https://rmarkdown.rstudio.com.
- Stephenson, Laura B; Harell, Allison; Rubenson, Daniel; Loewen, Peter John, 2020, '2019 Canadian Election Study - Online Survey',
           https://doi.org/10.7910/DVN/DUS88V, Harvard Dataverse, V1




